Manuscript 1 code

### 2024-09-20 manuscript_1_code.Rmd
fixed habitat levels to be in the original dataframe :tropical_fish:

not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")

@@

macro_species_cols <- macroalgae[,-which(names(macroalgae) %in% meta_cols)]

@@

cover_long <- left_join(cover_long_means,cover_long_se)

#make levels for habitat for plotting
cover_long$habitat <- factor(cover_long$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

cover_long$functional_group <- factor(cover_long$functional_group, 
                             levels = c("macroalgae", "coral", "turf","cca", "sand", "other"),
                             labels = c("Macroalgae", "Coral", "Turf", "CCA", "Sand", "Other"))
```

```{r simplified cover - remove}

# simplified_cover <- data.frame(
#   year = cover_df$year,
#   habitat = cover_df$habitat,
#   site = cover_df$site,
#   location = cover_df$location,
#   low_lying_algae = cover_df$cca + cover_df$turf + cover_df$cyanophyta,
#   coral = cover_df$coral,
#   macroalgae = cover_df$macroalgae
# )
# 
# simplified_cover_means <- simplified_cover %>%
#   group_by(year, habitat) %>%
#   summarise(macroalgae_mean = mean(macroalgae),
#             low_lying_algae_mean = mean(low_lying_algae),
#             coral_mean = mean(coral),
#             macroalgae_se = sd(macroalgae)/sqrt(n()),
#             low_lying_algae_se = sd(low_lying_algae)/sqrt(n()),
#             coral_se = sd(coral)/sqrt(n())
#             )
# 
# simplified_cover_long <- pivot_longer(
#   data = simplified_cover_means,
#   cols = 3:5,
#   names_to = "functional_group",
#   values_to = "percent_cover"
# )
# 
# simplified_cover_long_se <- pivot_longer(
#   data = simplified_cover_means,
#   cols = 6:8,
#   names_to = "functional_group",
#   values_to = "se"
# )
# simplified_cover_long$se <- simplified_cover_long_se$se

```{r}
figure_1_data <- subset(cover_long, functional_group != "Sand" & functional_group != "Other")
```

## plot
```{r figure 1, fig.height = 10, fig.width=14, warning=FALSE}
(figure_1 <- ggplot() +
  geom_point(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
  geom_errorbar(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group),
                width = 0, size = 1.3) +
   geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 1) +
  facet_wrap(~ habitat) +
  annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
  scale_colour_manual(values = functional_group_colours) +
  # scale_fill_manual(values = functional_group_colours) +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(cover_long$year)) +
   labs(y = "Percent Cover \n",
        x = "Year") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

@@

```{r richness factors}
quad_richness_df$habitat <- factor(quad_richness_df$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

@@

```{r richness factors}
site_richness_df$habitat <- factor(site_richness_df$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

@@

quad_evenness_df <- alpha_diversity_quad_macro_with_cover %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(evenness_mean = mean(evenness),
            evenness_se = se(evenness))

quad_evenness_df$habitat <- factor(quad_evenness_df$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

@@

## 2d Evenness @ Site by Year
### prep
```{r set up evenness}
site_evenness_df <- alpha_diversity_site_macro %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(evenness_mean = mean(evenness),
            evenness_se = se(evenness))

site_evenness_df$habitat <- factor(site_evenness_df$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

@@

alpha_diversity_quad_macro_with_cover$habitat <- factor(alpha_diversity_quad_macro_with_cover$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

filtered_cover_effects$habitat <- factor(filtered_cover_effects$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

@@

## Figure 3B quad level evenness versus cover
```{r COME BACK HERE figure 3b setup}
diversity_ranges_quad_evenness <- alpha_diversity_quad_macro_with_cover_evenness_mod %>%
  group_by(habitat) %>%
  summarise(min_evenness = min(evenness),
            max_evenness = max(evenness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

@@

alpha_diversity_quad_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_with_cover_evenness_mod$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

filtered_cover_effects_evenness$habitat <- factor(filtered_cover_effects_evenness$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

@@

alpha_diversity_site_macro_with_cover$habitat <- factor(alpha_diversity_site_macro_with_cover$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

filtered_site_cover_effects$habitat <- factor(filtered_site_cover_effects$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

@@

alpha_diversity_site_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_site_macro_with_cover_evenness_mod$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

@@

```{r figure 4a setup}

rich_stab_emm <- emmip(rich_stab_mod,
                        habitat ~ richness_mean,
                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects <- rich_stab_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)

@@

```{r figure 4c setup}

synch_stab_emm <- emmip(synch_stab_mod,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_stab_effects <- synch_stab_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)

@@

```{r figure 4d setup}

rich_stab_emm_site <- emmip(rich_stab_mod_site,
                        habitat ~ richness_mean,
                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_site <- rich_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)

@@

```{r figure 4f setup}

synch_stab_emm_site <- emmip(synch_stab_mod_site,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_stab_effects_site <- synch_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)

@@

```{r figure 4d setup}

even_stab_emm_site <- emmip(even_stab_mod_site,
                        habitat ~ evenness_mean,
                        at = list(evenness_mean = seq(0,1,0.01)), plotit = F, CIs = T)
#make levels for habitat for plotting

@@

#make levels for habitat for plotting
diversity_stability_synchrony_site$habitat <- factor(diversity_stability_synchrony_site$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

>>>>

#THIS IS THE BIG CHANGE HERE: 9/19
filtered_data$habitat <- factor(filtered_data$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

#DID NOT TOUCH BELOW:
not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")

@@

macro_species_cols <- macroalgae[,-which(names(macroalgae) %in% meta_cols)]
```

@@

cover_long <- left_join(cover_long_means,cover_long_se, by = join_by(year, habitat, functional_group))

#make levels of functional group for plotting
cover_long$functional_group <- factor(cover_long$functional_group, 
                             levels = c("macroalgae", "coral", "turf","cca", "sand", "other"),
                             labels = c("Macroalgae", "Coral", "Turf", "CCA", "Sand", "Other"))
```

```{r make figure1 data}
figure_1_data <- subset(cover_long, functional_group != "Sand" & functional_group != "Other")
```

## plot
```{r figure 1, fig.height = 10, fig.width=14, warning=FALSE}
(figure_1 <- ggplot() +
  geom_point(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
  geom_errorbar(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group),
                width = 0, size = 1.3) +
   geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
  facet_wrap(~ habitat) +
  annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
  scale_colour_manual(values = functional_group_colours) +
  # scale_fill_manual(values = functional_group_colours) +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(cover_long$year)) +
   labs(y = "Percent Cover \n",
        x = "Year") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

@@

@@

@@

quad_evenness_df <- alpha_diversity_quad_macro_with_cover %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(evenness_mean = mean(evenness),
            evenness_se = se(evenness))
```

@@

## 2d Evenness @ Site by Year
### prep
```{r set up evenness}
site_evenness_df <- alpha_diversity_site_macro %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(evenness_mean = mean(evenness),
            evenness_se = se(evenness))
```

@@

# alpha_diversity_quad_macro_with_cover$habitat <- factor(alpha_diversity_quad_macro_with_cover$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
# 
# 
# filtered_cover_effects$habitat <- factor(filtered_cover_effects$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

@@

## Figure 3B quad level evenness versus cover
```{r figure 3b setup}
diversity_ranges_quad_evenness <- alpha_diversity_quad_macro_with_cover_evenness_mod %>%
  group_by(habitat) %>%
  summarise(min_evenness = min(evenness),
            max_evenness = max(evenness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

@@

# alpha_diversity_quad_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_with_cover_evenness_mod$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
# 
# 
# filtered_cover_effects_evenness$habitat <- factor(filtered_cover_effects_evenness$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

@@

# alpha_diversity_site_macro_with_cover$habitat <- factor(alpha_diversity_site_macro_with_cover$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
# 
# 
# filtered_site_cover_effects$habitat <- factor(filtered_site_cover_effects$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

@@

# alpha_diversity_site_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_site_macro_with_cover_evenness_mod$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
# 
# 
# 
```

@@

```{r figure 4a setup}
rich_stab_emm <- emmip(rich_stab_mod,
                        habitat ~ richness_mean,
                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects <- rich_stab_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)
```

@@

```{r figure 4c setup}
synch_stab_emm <- emmip(synch_stab_mod,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_stab_effects <- synch_stab_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
```

@@

```{r figure 4d setup}
rich_stab_emm_site <- emmip(rich_stab_mod_site,
                        habitat ~ richness_mean,
                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_site <- rich_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)
```

@@

```{r figure 4f setup}
synch_stab_emm_site <- emmip(synch_stab_mod_site,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_stab_effects_site <- synch_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
```

@@

```{r figure 4d setup}
even_stab_emm_site <- emmip(even_stab_mod_site,
                        habitat ~ evenness_mean,
                        at = list(evenness_mean = seq(0,1,0.01)), plotit = F, CIs = T)
#make levels for habitat for plotting

@@

#make levels for habitat for plotting
# diversity_stability_synchrony_site$habitat <- factor(diversity_stability_synchrony_site$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```
